---
title: "Building a neural network for MNIST data"
author: "Chad Allison | 29 November 2022"
output: github_document
---

___

Here, I will build a neural network model to identify handwritten digits using the [MNIST dataset](https://www.kaggle.com/c/digit-recognizer/overview) which I have obtained from Kaggle.

___

### Importing libraries and setting some quality of life options

```{r message = F, warning = F}
library(tidyverse)
library(keras)
library(tensorflow)
library(yardstick)
library(tidymodels)

knitr::opts_chunk$set(message = F, warning = F)
options(scipen = 999)
theme_set(theme_classic())
```

___

### Importing the data

```{r}
mnist = read_csv("mnist_data.csv", col_types = cols())
head(mnist)
```

The data has 42,000 observations with 785 variables. The first variable, `label` is our outcome variable, while the rest of the variables are the brightness of the pixel indicated by the variable name.

___

### Cross validation

We will split the data into training and testing data, electing to use 80% as training data and the remaining 20% as testing data. We will also identify `label` as our outcome variable in order to preserve balance between classes.

```{r}
split = initial_split(mnist, prop = 0.8, strata = "label")
train = training(split)
test = testing(split)

# creating a `set` variable to combine the two and create visual
train$set = "train"
test$set = "test"

rbind(train, test) |>
  mutate(label = factor(label)) |>
  group_by(set) |>
  count(label) |>
  mutate(prop = ifelse(set == "train", round(n / nrow(train), 3), round(n / nrow(test), 3))) |>
  ggplot(aes(label, prop)) +
  geom_col(aes(fill = set), position = "dodge") +
  scale_fill_manual(values = c("#CAA2D0", "#90BD90")) +
  labs(title = "class balance between testing and training data") +
  theme(plot.title = element_text(hjust = 0.5))

# removing the `set` variable for later use
train = train |> select(-set)
test = test |> select(-set)
```















































































